---
title: "Three arm simulation"
subtitle: "Variations on a theme"
author: "Keaven M. Anderson"
date: "2023-01-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, parallelly.fork.enable = TRUE)
```

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(tibble)
library(gt)
library(survival)
library(mvtnorm)
library(gsDesign)
library(gsDesign2)
library(ggplot2)
library(cowplot)
library(simtrial)
```



# Introduction

This is to demonstrate a simulation with more than 2 arms and with different dropout rates in each arm using `simtrial::simPWSurv()`.
The simplest case is for 3 arms and a single stratum.
There are 3 parts to document:

1) Assumptions for enrollment and time-to-event distribution by treatment group.
2) Generating data for multiple treatment groups including differential censoring rates.
3) Testing for advantage of 2 experimental treatments over control using different multiplicity strategies.

```{r}
scen <- 1 # Select scenario to simulate
```

# Assumptions for example

We lay out assumptions. We assume a single stratum with blocks of 6 (2:2:2) randomization (Control:Ex1:Ex2).
We assume a constant enrollment rate until the targeted sample size ('n') is reached.

```{r}
# 3 treatment groups
tx <- c("Control", "Ex1", "Ex2")
# Blocks of 5 with 1:2:2 randomization
block <- c("Control", "Control", "Ex1", "Ex1", "Ex2", "Ex2")
```

For enrollment, we assume ramp-up for 4 months followed by 10 months of study state enrollment for a total enrollment duration of 14 months.

```{r}
enroll_rate <- tibble(stratum = "All", duration = c(2, 2, 10), rate = c(3, 6, 9))
enroll_duration <- sum(enroll_rate$duration)
```

Now we set failure rates for each treatment group for 6 different scenarios. 
We can have different piecewise exponential distribution rates for each treatment group.

## Failure rate scenarios

```{r, echo=FALSE, message = FALSE}
tdp <- 40

failure_rates <- tribble(
  ~Scenario, ~Name, ~Period, ~Treatment, ~Median, ~HR, ~Duration,
  1, "Design", 1, "Control", 14, 1, tdp,
  1, "Design", 1, "Ex1", 14, 1, 4,
  1, "Design", 2, "Ex1", 14, .6, tdp - 4,
  1, "Design", 1, "Ex2", 14, 1.2, 6,
  1, "Design", 2, "Ex2", 14, .55, tdp - 6,
  2, "NULL", 1, "Control", 14, 1, tdp,
  2, "NULL", 1, "Ex1", 14, 1, tdp,
  2, "NULL", 1, "Ex2", 14, 1, tdp,
  3, "Partial NULL", 1, "Control", 14, 1, tdp,
  3, "Partial NULL", 1, "Ex1", 14, 1, 4,
  3, "Partial NULL", 2, "Ex1", 14, .6, tdp - 4,
  3, "Partial NULL", 1, "Ex2", 14, 1, tdp,
  4, "Longer TTE", 1, "Control", 17, 1, tdp,
  4, "Longer TTE", 1, "Ex1", 17, 1, 4,
  4, "Longer TTE", 2, "Ex1", 17, .6, tdp - 4,
  4, "Longer TTE", 1, "Ex2", 17, 1.2, 6,
  4, "Longer TTE", 2, "Ex2", 17, .55, tdp - 6,
  5, "Shorter TTE", 1, "Control", 11, 1, tdp,
  5, "Shorter TTE", 1, "Ex1", 11, 1, 4,
  5, "Shorter TTE", 2, "Ex1", 11, .6, tdp - 4,
  5, "Shorter TTE", 1, "Ex2", 11, 1.2, 6,
  5, "Shorter TTE", 2, "Ex2", 11, .55, tdp - 6,
  6, "PH", 1, "Control", 14, 1, tdp,
  6, "PH", 1, "Ex1", 14, .68, tdp,
  6, "PH", 1, "Ex2", 14, .8, tdp
)

xx <-
  failure_rates %>%
  group_by(Scenario, Treatment) %>%
  summarize(
    Name = Name, Stratum = "All", rate = log(2) / Median * HR, Time = cumsum(Duration),
    "Hazard" = cumsum(rate * Duration), Survival = exp(-Hazard)
  ) %>%
  ungroup() %>%
  select(c(Scenario, "Name", "Treatment", "Time", "Survival"))

xxsurv0 <- xx %>%
  group_by(Scenario, Treatment) %>%
  filter(row_number() == 1) %>%
  mutate(Time = 0, Survival = 1)

ggplot(rbind(xxsurv0, xx), aes(x = Time, y = Survival, col = Treatment)) +
  geom_line() +
  geom_point() +
  facet_wrap(facet = vars(Name), ncol = 3) +
  scale_x_continuous(breaks = seq(0, 40, 8)) +
  scale_y_log10(breaks = c(.125, .25, .5, .75)) +
  theme_half_open(12) +
  background_grid(minor = "none")

fr_sim <-
  failure_rates %>%
  transmute(
    Scenario = Scenario, Name = Name, Stratum = "All", period = Period, Treatment = Treatment,
    HR = HR, duration = Duration, rate = log(2) / Median * HR
  )
dropoutRates <- tibble(
  Stratum = "All", period = 1,
  Treatment = c("Control", "Ex1", "Ex2"), duration = 100, rate = .001
)
```

# Design: using gsDesign2::gs_design_ahr()

Delayed effect with HR = 0.6 after 4 months, 
Control: exponential time to failure, median = 14 months

```{r, echo=FALSE}
design <- gs_design_ahr(
  enroll_rate = tibble(stratum = "All", duration = c(2, 2, 10), rate = c(3, 6, 9)),
  fail_rate = failure_rates %>%
    filter(Scenario == 1, Treatment == "Ex1") %>%
    mutate(
      stratum = "All", duration = Duration,
      fail_rate = log(2) / Median, hr = HR,
      dropout_rate = .001
    ),
  alpha = 0.0125,
  beta = 0.2,
  info_frac = c(.4, .8, 1),
  analysis_time = 36,
  ratio = 1,
  binding = FALSE,
  upper = gs_spending_bound,
  upar = list(sf = sfLDOF, sfupar = 1, total_spend = 0.0125, param = NULL, timing = NULL),
  lower = gs_b,
  lpar = c(0, -Inf, -Inf),
  r = 18,
  tol = 1e-06
)

N <- last(design$analysis$n)
enroll_rate <- design$enroll_rate  %>% mutate(rate = round(2 * ceiling(N / 2) / N * rate),0) 
event <- ceiling(design$analysis$event)

designi <- gs_power_ahr(
  enroll_rate = enroll_rate,
  fail_rate = design$fail_rate,
  event = event,
  analysis_time = c(.1, .2, .3),
  ratio = 1,
  binding = FALSE,
  upper = gs_spending_bound,
  upar = list(sf = sfLDOF, sfupar = 1, total_spend = 0.0125, param = NULL, timing = NULL),
  lower = gs_b,
  lpar = c(0, -Inf, -Inf),
  r = 18,
  tol = 1e-06
)

designi %>%
  summary() %>%
  as_gt() %>%
  tab_header("Planned design for 2 arms, alpha=0.0125, 4 month effect delay",
    subtitle = "Increase sample size by 50% for 3-arm trial"
  )

N <- last(designi$analysis$n)
events <- designi$analysis$event
analysis_times <- designi$analysis$time
```

# Simulation

Now we simulate `r N` observations as suggested in the above design. 
This is the raw dataset which will subsequently require cutting off for analysis.
We show the first 2 blocks (12 observations) sorted by the first of time-to-event and censoring.

```{r}
gendat <- function(scen = 1,
                   Nobs,
                   rand_block,
                   enrollRates,
                   failRates,
                   dropoutRates) {
  ans <- simPWSurv(
    n = ceiling(Nobs),
    strata = tibble(Stratum = "All", p = 1),
    block = rand_block,
    enrollRates = enrollRates,
    failRates = failRates,
    dropoutRates = dropoutRates
  )
  return(ans)
}

raw_3_arm_data <- gendat(
  scen = 1,
  Nobs = N * 3 / 2,
  rand_block = block,
  enrollRates = designi$enroll_rate %>% mutate(rate = rate * 1.5),
  failRates = fr_sim %>% filter(Scenario == scen),
  dropoutRates = dropoutRates
)
raw_3_arm_data %>%
  ungroup() %>%
  head(n = 12) %>%
  arrange(enrollTime) %>%
  gt() %>%
  tab_header(title = "First 12 simulated observations", subtitle = "2 blocks of 6") %>%
  fmt_number(columns = c(2, 4:6), decimals = 2)
```

Now we find dates for analysis cuts based on planned date, pairwise comparison of Ex1 versus control, and pairwise comparison of Ex2 versus control. 

```{r}
get_cut_dates <- function(x, # raw data simulated by gendat()
                          planned_design # designi # planed design
) {
  # Planned date
  analysis_cut_dates <- tibble(
    Analysis = planned_design$analysis$analysis,
    planned_date = planned_design$analysis$time
  )

  # Ex1 versus control
  ex1controldat <- x %>% filter(Treatment != "Ex2")
  dt <- rep(0, length(planned_design$analysis$analysis))
  for (i in planned_design$analysis$analysis) {
    dt[i] <- ex1controldat %>% getCutDateForCount(count = planned_design$analysis$event[i])
  }
  analysis_cut_dates$Ex1Count <- dt

  # Ex2 versus control
  ex2controldat <- x %>% filter(Treatment != "Ex1")
  dt <- rep(0, length(planned_design$analysis$analysis))
  for (i in planned_design$analysis$analysis) {
    dt[i] <- ex2controldat %>% getCutDateForCount(count = planned_design$analysis$event[i])
  }
  analysis_cut_dates$Ex2Count <- dt

  # Get the max date of the above pairwise comparison
  analysis_cut_dates <- analysis_cut_dates %>% mutate(maxEx1Ex2 = pmax(Ex1Count, Ex2Count))
  method1 <- analysis_cut_dates %>% transmute(Method = 1, Analysis = Analysis, Date = planned_date)
  method2 <- analysis_cut_dates %>% transmute(Method = 2, Analysis = Analysis, Date = maxEx1Ex2)
  method1$Date[3] <- max(method1$Date[3], method2$Date[3])

  return(rbind(method1, method2))
}

analysis_cut_dates <- get_cut_dates(x = raw_3_arm_data, planned_design = designi)
analysis_cut_dates
```

# Analysis

We examine event counts and sample size by treatment group. We see that each of the experimental treatment groups has a smaller proportion with events than the control group.

```{r}
doAnalysis <- function(x = raw_3_arm_data, # simulated raw data by gendat()
                       cut_date_tbl, # cut date per analysis per method
                       imethod = 1, # index of method, either 1 or 2
                       ianalysis = 1, # index of analysis
                       sim = 1, # index of simulation
                       scen = 1 # index of scenario
) {
  # Get cut date
  cut_date <- (cut_date_tbl %>% filter(Method == imethod))$Date[ianalysis]

  # Cut the simulated data by the above cut date
  analysis_data <- x %>%
    cutData(cut_date) %>%
    mutate(
      ExAny = 1 * (Treatment != "Control"), # For combined treatment groups vs control
      dr = 1 * (Treatment == "Ex1") + 1.5 * (Treatment == "Ex2")
    )

  # Get events by treatment
  results_counts <- as.data.frame(t(table(analysis_data$Treatment, analysis_data$event)[, 2]))

  # Pairwise comparisons
  Z1 <- (analysis_data %>%
    filter(Treatment != "Ex2") %>%
    tensurv(txval = "Ex1") %>%
    tenFH(rg = tibble(rho = 0, gamma = 0)))$Z

  Z2 <- (analysis_data %>%
    filter(Treatment != "Ex1") %>%
    tensurv(txval = "Ex2") %>%
    tenFH(rg = tibble(rho = 0, gamma = 0)))$Z

  # Combined treatments vs control
  ZEx <- (analysis_data %>%
    mutate(Treatment = ExAny) %>%
    tensurv(txval = 1) %>%
    tenFH(rg = tibble::tibble(rho = 0, gamma = 0)))$Z

  # Wald test for dose-response
  xx <- coxph(Surv(tte, event) ~ dr, data = analysis_data)
  Zdr <- as.numeric(xx$coefficients / sqrt(xx$var))

  # Output results
  results_other <- tibble(
    Scenario = scen, sim = sim, Method = imethod, Analysis = ianalysis, cut_date = cut_date,
    Z1 = -Z1, Z2 = -Z2, ZEx = -ZEx, Zdr = -Zdr
  )

  return(cbind(results_other, results_counts))
}
```

Now we proceed to do the entire simulation and save the analysis results as summary data.
First we look at how many cores are available.

```{r}
parallelly::availableCores()
```

```{r}
# Check whether or not forked processing ("multicore") is "supported"
# Running this under RStudio will return FALSE, even if not under Windows
parallelly::supportsMulticore()
```

```{r, message=FALSE, warning=FALSE}
library(doFuture)
library(doRNG)
plan("multisession", workers = parallelly::availableCores())


```

```{r}
# Parallel setup
library(future.batchtools)
library(doFuture)
options(future.batchtools.workers=2)
plan(batchtools_slurm, template=batchtools::findTemplateFile("slurm-simple"), resources = list(ncpus=1, memory = "1g", walltime=120))
```



```{r}
start_time <- Sys.time()
nsim <- 5
# List for 2 methods times number of simulations
# Each list item will have 3 analysis rows
# results <- vector("list", length = nsim * 6)
set.seed(1)
results <- foreach(
  sim = seq_len(nsim),
  .combine = "rbind",
  .errorhandling = "pass",
  .options.future = list(seed = TRUE)
) %dofuture% {
  # Generate raw data
  raw_data <- gendat(
    scen = scen, Nobs = N * 3 / 2,
    rand_block = block,
    enrollRates = designi$enroll_rate %>% mutate(rate = rate * 1.5),
    failRates = fr_sim %>% filter(Scenario == scen),
    dropoutRates = dropoutRates
  )

  # TEST to see if data generated differs
  # return(raw_data[1,])
  analysis_cut_dates <- get_cut_dates(x = raw_data, planned_design = designi)

  # Test to see of get_cut_dates is working
  # return(analysis_cut_dates)
  do.call(
    rbind,
    lapply(1:3, function(y) {
      rbind(
        doAnalysis(x = raw_data, cut_date_tbl = analysis_cut_dates, sim = sim, imethod = 1, ianalysis = y),
        doAnalysis(x = raw_data, cut_date_tbl = analysis_cut_dates, sim = sim, imethod = 2, ianalysis = y)
      )
    })
  )
}
end_time <- Sys.time()
sim_duration <- end_time - start_time
sim_duration
```

```{r}
head(results, n = 18) %>% gt()
```

```{r}
save(list = c("results", "nsim", "sim_duration"), file = "../Scenario1_10000x.Rdata")
```
