---
title: "cmdstanr"
output: html_notebook
---

This is an example on how to use [cmdstanr](https://mc-stan.org/cmdstanr/).

# Prerequisites (one-time setup)

First of all, you need to install `cmdstan`. We are making use of all the cores we have available via the `parallelly` package. Compiling C++ code is inherently slow so let's make sure we use all available resources. Additionally we are specifying a specific `cmdstan` version for both good practice and increased reproducibility. 

```{r message=FALSE, warning=FALSE}
library(cmdstanr)
cmdstanr::install_cmdstan(
  version = "2.34.1", 
  cores = parallelly::availableCores()
  )
```

Please note you always can rebuild cmdstan via the `rebuild_cmdstan()` function. 

# Example

## Creating a model

We will download the `bernoulli_logit_glm` model from the `cmdstanr` vignette and save it as `model.stan`. 

```{r}
write(
  RCurl::getURI(
    "https://raw.githubusercontent.com/stan-dev/cmdstanr/master/vignettes/articles-online-only/opencl-files/bernoulli_logit_glm.stan"
    ),
  "model.stan")
```

## Data creation

```{r}
n <- 25000
k <- 20
X <- matrix(rnorm(n * k), ncol = k)
y <- rbinom(n, size = 1, prob = plogis(3 * X[,1] - 2 * X[,2] + 1))
mdata <- list(k = k, n = n, y = y, X = X)
```

## Compile the model 

```{r}
mod_cpu <- cmdstanr::cmdstan_model("model.stan")
```

## Run the model

```{r}
system.time(fit_cpu <- mod_cpu$sample(data = mdata, chains = 4, parallel_chains = 4, refresh = 0))
```

# Appendix

## Working with GPUs 

`cmdstan` uses OpenCL to make use of a GPU. We will need to build `cmdstan` and link against OpenCL. 

```{r}
cpp_options = list(
  "LDFLAGS+= -lOpenCL"
)

cmdstan_dir="~/.cmdstan-gpu"
dir.create(cmdstan_dir)
cmdstanr::install_cmdstan(version="2.34.1",
                          dir=cmdstan_dir,
                          cpp_options=cpp_options,
                          cores=parallelly::availableCores())
```

For each `cmdstan` execution leveraging GPUs we now have to point to this `cmdstan` install

```{r}
cmdstanr::set_cmdstan_path(paste(cmdstan_dir,"cmdstan-2.34.1",sep="/"))
```
